---
title: "Cross-Modal Head Nod Analysis - Interactive Notebook"
output: 
  html_notebook:
    toc: true
    toc_float: true
    theme: cosmo
    highlight: tango
---

# Cross-Modal Head Nod Analysis

This interactive notebook provides an overview of the cross-modal head nod kinematic analysis comparing German Sign Language (DGS), German, Russian Sign Language (RSL), and Russian.

## Quick Start

Run the following code chunk to initialize the analysis environment:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Check if we're in Binder environment
if (Sys.getenv("BINDER_SERVICE_HOST") != "") {
  cat("🐳 Running in Binder environment\n")
  # Install packages if needed
  if (!file.exists("install.R.binder")) {
    source("install.R")
  } else {
    source("install.R.binder")
  }
} else {
  cat("💻 Running in local environment\n")
  source("install.R")
}

# Load configuration
source("config.R")
```

## Repository Overview

This repository contains:

- **Data**: Kinematic measurements from 4 languages across 4 tiers (form, function, turn, functionturn)
- **Scripts**: 8 R analysis scripts for comprehensive statistical analysis
- **Configuration**: Portable configuration system (`config.R`)
- **Testing**: Comprehensive validation framework (`test_repository.R`)

## Key Variables

The analysis focuses on three kinematic properties:

```{r variables}
cat("📊 Kinematic Variables:\n")
for (i in seq_along(CONFIG$KINEMATIC_VARS)) {
  cat(sprintf("   %d. %s\n", i, CONFIG$KINEMATIC_VARS[i]))
}

cat("\n🌍 Languages:\n")
for (i in seq_along(CONFIG$LANGUAGES)) {
  cat(sprintf("   %d. %s\n", i, CONFIG$LANGUAGES[i]))
}
```

## Data Overview

```{r data-overview}
# Load main dataset
if (file.exists(CONFIG$FUNCTION_DATA)) {
  data <- read.csv(CONFIG$FUNCTION_DATA)
  
  cat("📁 Function Data Overview:\n")
  cat(sprintf("   Rows: %d\n", nrow(data)))
  cat(sprintf("   Columns: %d\n", ncol(data)))
  cat(sprintf("   Languages: %s\n", paste(unique(data$language), collapse = ", ")))
  cat(sprintf("   Functions: %s\n", paste(unique(data$Label), collapse = ", ")))
  
  # Show first few rows
  cat("\n📋 Sample Data:\n")
  print(head(data[, c("language", "Label", CONFIG$KINEMATIC_VARS)], 10))
} else {
  cat("❌ Data files not found. Please ensure data/ directory contains the required CSV files.\n")
}
```

## Quick Analysis Example

```{r quick-analysis}
library(ggplot2)
library(dplyr)

if (exists("data") && nrow(data) > 0) {
  # Create a quick visualization
  plot_data <- data %>%
    filter(!is.na(length..seconds.)) %>%
    mutate(modality = ifelse(grepl("DGS|RSL", language), "Signed", "Spoken"))
  
  p <- ggplot(plot_data, aes(x = modality, y = length..seconds., fill = modality)) +
    geom_boxplot(alpha = 0.7) +
    scale_fill_manual(values = c("Signed" = "#2E86AB", "Spoken" = "#A23B72")) +
    labs(
      title = "Head Nod Duration by Modality",
      x = "Modality",
      y = "Duration (seconds)",
      fill = "Modality"
    ) +
    theme_minimal() +
    theme(legend.position = "none")
  
  print(p)
  
  # Quick statistics
  summary_stats <- plot_data %>%
    group_by(modality) %>%
    summarise(
      n = n(),
      mean_duration = mean(length..seconds., na.rm = TRUE),
      sd_duration = sd(length..seconds., na.rm = TRUE),
      .groups = 'drop'
    )
  
  cat("\n📈 Quick Statistics:\n")
  print(summary_stats)
}
```

## Running Full Analysis

To run the complete analysis pipeline:

```{r full-analysis, eval=FALSE}
# 1. Data Preparation
source("scripts/01_data_preparation.R")

# 2. Comprehensive Analysis  
source("scripts/02_comprehensive_analysis.R")

# 3. Advanced Statistics
source("scripts/03_advanced_statistics.R")

# 4. Refined Analyses
source("scripts/04_refined_analyses.R")

# 5. Publication Figures
source("scripts/05_publication_figures.R")
```

## Repository Validation

Test the complete repository functionality:

```{r validation, eval=FALSE}
# Run comprehensive validation
source("test_repository.R")
```

## Links and Resources

- **GitHub Repository**: [jobschepens/cross-modal-head-nod-analysis](https://github.com/jobschepens/cross-modal-head-nod-analysis)
- **Launch in Binder**: [![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/jobschepens/cross-modal-head-nod-analysis/HEAD?urlpath=rstudio)
- **Documentation**: See `README.md` for detailed setup instructions

## Interactive Environment

This notebook is running in an interactive environment. You can:

- ✅ Modify and run code chunks
- ✅ Explore data interactively  
- ✅ Generate custom visualizations
- ✅ Run statistical analyses
- ✅ Export results and figures

Happy analyzing! 🎉
